SHORT: Longitudinal Study of Software Environments Produced by Dockerfiles from Research Artifacts: Initial Design

### 2025-05-20 src/main.tex
submitted version
aspell on the text

One difficult aspect of creating a truly reproducible artifact is to properly control the software environment: packages and libraries versions.
Several works already shed light on the possible variability of results given a badly captured software environment~\cite{mytkowicz_producing_nodate, sokolowski2024impact}.
To address this issue, the conference \adae\ committees have advised authors to provide containers or virtual machines to share their artifacts in order to ``freeze'' their software environment~\cite{eurosys25, aetips, csartifacts}.
In practice, about 12\% of the artifacts use such tools~\cite{guilloteau2024longevity}.
%However, authors generate their artifact right before the artifact submission~\cite{guilloteau2024longevity, guilloteau2024frustrations}.
In the case of usual virtualization techniques to generate containers or virtual machines, authors have to write the ``recipe'' (for example, \df\ for Docker containers) made of a list of commands to execute from a ``base image'' (the \texttt{FROM} instruction in \df s).
This approach has several limitations.
First, these base images can vary, either because of bad practices from the users (\eg\ using the \texttt{latest} version of the images) or from the administrators of the image (\eg\ pushing a different image to a previously defined tag).
Administrators can also remove these base images, making all recipes un-buildable~\cite{nvidia_cuda_lifetime}.
Another limitation of this approach is the heavy dependency on non-reproducible tools such as classical package managers (\eg\ \texttt{apt}, \texttt{dpkg}, \texttt{yum}).
Hence, rebuilding a recipe of a container or a virtual machine can yield different results from the authors' image, and thus endangers the reproducibility of the results.
One would be for the authors to store and then provide the built container itself (\ie tarball) on some long-term storage (\eg\ Zenodo~\cite{zenodo}).
But this comes with the downside of requiring more storage space than just storing the recipe (binary vs. text file), and thus has some \textbf{severe scalability and sustainability issues in terms of longevity}~\cite{monroe2023preservation, guilloteau2024longevity}.
Moreover, storing the built binary image exhibits limitations when needed to \emph{inspect} the actual content of the image, and when a modification of the software environment capture by this image is needed~\cite{mercier2018considering}.
Having the opportunity to extend previous scientific work is a crucial aspect in any scientific field.

@@

With \ecg, a \df\ from an artifact is represented as a Nickel configuration file~\cite{nickel}.
Nickel is a configuration language with a functional paradigm.
One of Nickel's features is the possibility to \emph{verify} that a configuration file is valid with respect to a \emph{contract}.
The Nickel file representing the artifact is then \emph{verified} against the contract~\cite{ecg_contract} and translated to a JSON file given as input to \ecg.
Listing~\ref{lst:nickel} shows an example of a Nickel representation.
The contract is versioned (line 2 of Listing~\ref{lst:nickel}) to take into account potential future changes to the descriptions and the contract.

@@

For example, authors can link to a Zenodo archive that will \emph{always} point to the latest artifact version.% of an artifact.
Hence, the content of the artifact behind this link can change over time.

@@

@@

The workflow of \ecg\ is shown in Figure~\ref{fig:workflow}.
The workflow itself~\cite{ecg_workflow} is governed by a Snakemake~\cite{koster2012snakemake} description, and its software environment~\cite{ecg_nix} is managed with Nix~\cite{dolstra_nix_2004}.

@@

\begin{enumerate}
\item For each conference with an artifact evaluation process, we examine all artifacts and consider only those that share their software environment using a \df.
\item From the artifact and the \df we generate the associated Nickel file (Listing~\ref{lst:nickel}) which contains information such as the artifact's location, how to extract the \df\ from the artifact and the package managers used. 
%where is the artifact stored? how to extract the \df\ out from the artifact? which package managers are used?
\item Once all Nickel descriptions for the artifacts of a conference have been completed, we use \ecg\ to construct the download, build, and extract information from each of them.
\item Prior to building each container, the local Docker cache is cleared on these machines.
\item We will execute \ecg\ on each artifact every month, as close to the beginning of the month as possible, for a full year (\ie\ 13 months).
\end{enumerate}

\paragraph{Details}
\begin{itemize}
\item We ensure that an empty local Docker cache is used before building the containers.
\item Each container is allocated a 10-minute time frame to build.
\item \ecg\ is executed on the x86-64 machines of the \grid\ platform~\cite{grid5000} on the \texttt{dahu} cluster of the Grenoble site.
\end{itemize}

@@

The community may want to adopt solutions \emph{tailored for reproducibility} of software environment (\ie\ Nix~\cite{dolstra_nix_2004} and Guix~\cite{courtes_functional_2013}), where authors can share a textual representation of their software environment which could be precisely reproduced and with longevity guarantees~\cite{courtes2024source}.
Moreover, Nix and Guix allow the precise introduction of variation in the software environment, which is a crucial feature (not present in container technologies) in order to build upon previous artifact for future publications~\cite{mercier2018considering}.

@@

\section*{Acknowledgments}

@@

\newpage
\newpage
%\bibliographystyle{sty/ACM-Reference-Format}
%\bibliography{references}
\printbibliography

>>>>

One difficult aspect of creating a truly reproducible artifact is to properly control the software environment: packages and libraries versions.
Several works already shed light on the possible variability of experimental results given a badly captured software environment~\cite{mytkowicz_producing_nodate, sokolowski2024impact}.
To address this issue, the conference \adae\ committees have advised authors to provide containers or virtual machines to share their artifacts in order to ``freeze'' their software environment~\cite{eurosys25, aetips, csartifacts}.
In practice, about 12\% of the artifacts use such tools~\cite{guilloteau2024longevity}.
%However, authors generate their artifact right before the artifact submission~\cite{guilloteau2024longevity, guilloteau2024frustrations}.
In the case of usual virtualization techniques to generate containers or virtual machines, authors have to write the ``recipe'' (for example, \df\ for Docker containers) made of a list of commands to execute from a ``base image'' (the \texttt{FROM} instruction in \df s).
This approach has several limitations.
First, these base images can vary, either because of bad practices from the users (\eg\ using the \texttt{latest} version of the images) or from the administrators of the image (\eg\ pushing a different image to a previously defined tag).
Administrators can also remove these base images, making all recipes un-buildable~\cite{nvidia_cuda_lifetime}.
Another limitation of this approach is the heavy dependency on non-reproducible tools such as classical package managers (\eg\ \texttt{apt}, \texttt{dpkg}, \texttt{yum}).
Hence, rebuilding a recipe of a container or a virtual machine can yield different results from the authors' image, and thus endangers the reproducibility of the results.
One would be for the authors to store and then provide the built container itself (\ie tarball) on some long-term storage (\eg\ Zenodo~\cite{zenodo}).
But this comes with the downside of requiring more storage space than just storing the recipe (binary vs. text file), and thus has some \textbf{severe scalability and sustainability issues in terms of longevity}~\cite{monroe2023preservation, guilloteau2024longevity}.
Moreover, storing the built binary image exhibits limitations when needed to \emph{inspect} the actual content of the image, and when a modification of the software environment capture by this image is needed~\cite{mercier2018considering}.
Having the opportunity to extend previous scientific work is a crucial aspect in any scientific field.
To the best of our knowledge, no previous work has exhibited the variations in software environment build from \df s through time.

@@

@@

For example, authors can link to a Zenodo archive that will \emph{always} point to the latest artifact version.
Hence, the content of the artifact behind this link can change over time.

@@

The information captured by \ecg\ cannot be automatically extracted from a \df.
In some cases, authors of artifact can have shell scripts installing packages (\eg\ calling \texttt{apt install}), which would be invisible at the \df\ level.
With \ecg, a \df\ from an artifact is represented as a Nickel configuration file~\cite{nickel}, which is similar to JSON or YAML.
%Nickel is a configuration language with a functional paradigm.
One of Nickel's features is the possibility to \emph{verify} that a configuration file is valid with respect to a \emph{contract}.
The Nickel file representing the artifact is then \emph{verified} against the contract~\cite{ecg_contract} and translated to a JSON file given as input to \ecg.
Listing~\ref{lst:nickel} shows an example of a Nickel representation.
The contract is versioned (line 2 of Listing~\ref{lst:nickel}) to take into account potential future changes to the descriptions and the contract.
If the Nickel file does not respect the contract, an error is reported to the user.

@@

The workflow of \ecg\ is shown in Figure~\ref{fig:workflow}.
The workflow itself~\cite{ecg_workflow} is governed by the Snakemake workflow manager~\cite{koster2012snakemake}, and its software environment~\cite{ecg_nix} is managed by the Nix functional package manager~\cite{dolstra_nix_2004}.
These tools provide reproducibility and robustness to the workflow setup to ensure that we will be able to execute the exact same workflow in the exact same conditions for all the future builds of the artifacts.

@@

\begin{enumerate}
\item For each conference with an artifact evaluation process, we examine all artifacts and consider only those that share their software environment using a \df.
\item From the artifact and the \df\ we generate the associated Nickel file (Listing~\ref{lst:nickel}) which contains information such as the artifact's location, how to extract the \df\ from the artifact and the package managers used. 
%where is the artifact stored? how to extract the \df\ out from the artifact? which package managers are used?
\item Once all Nickel descriptions for the artifacts of a conference have been completed, we use \ecg\ to construct the download, build, and extract information from each of them.
\item Prior to building each container, the local Docker cache is cleared on these machines.
\item We will execute \ecg\ on each artifact every month, as close to the beginning of the month as possible, for a full year (\ie\ 13 months).
\end{enumerate}

\paragraph{Details}
\begin{itemize}
\item We ensure that an empty local Docker cache is used before building the containers.
\item Each container is allocated a 10-minute time frame to build.
\item \ecg\ is executed on the x86-64 machines of the \grid\ platform~\cite{grid5000} on the \texttt{dahu} cluster of the Grenoble site.
\item We do not try to reproduce the experiment results from the artifacts, but only to build the container. The impact of software environment variations on the experiments results has been studied for example in \cite{mytkowicz_producing_nodate, sokolowski2024impact}.
\end{itemize}

@@

The community may want to adopt instead solutions \emph{tailored for long-term reproducibility} of software environment~\cite{malka2024reproducibility} (\ie\ Nix~\cite{dolstra_nix_2004} and Guix~\cite{courtes_functional_2013}), where authors can share a textual representation of their software environment which could be precisely reproduced and with longevity guarantees~\cite{courtes2024source}.
Moreover, Nix and Guix allow the precise introduction of variation in the software environment, which is a crucial feature (not present in container technologies) in order to build upon previous artifact for future publications~\cite{mercier2018considering}.

@@

\newpage
\section*{Acknowledgments}

@@

% \newpage
% \newpage
%\bibliographystyle{sty/ACM-Reference-Format}
%\bibliography{references}
\printbibliography
